Article Title

### 2013-09-23 sections/Estimators.tex
Kreso's comments

The hyperparameters $\hat\eta$ and $\hat\lambda$ must be estimated from the data.  Under the MSE loss function $\mathcal L_e$ (\ref{eq:MSE}), the optimal values of $\hat\eta$ and $\hat\lambda$ can be estimated analytically \cite{Ledoit:2004,Schafer:2005,Schaefer:2010}. However, these estimates are no longer optimal under the Guassian loss $\mathcal L_g$. \TODO{I have tested this and there was a substantial difference in both synthetic and empirical data}  When an analytical solution for optimal hyperparameters is not available, the optimal values can be found by cross validation, entirely within the the training sample. \TODO{explain nested cross-validation in more detail?}

@@

\subsection*{Estimator B: Shrinkage toward a factor model}
In estimator B, the target is a factor model, composed as the sum of a low-rank component $\hat L_{\hat d}\hat L_{\hat d}^\T$ and a diagonal matrix $\hat \Psi$:
\begin{equation}
\hat T_{\hat d} = \hat L_{\hat d} \hat L_{\hat d}^\T + \hat\Psi
\end{equation}
Here $\hat L_{\hat d}$ is an $p\times\hat d$ matrix and $\hat\Psi$ is diagonal.

@@

Just as in Estimator A, Estimator B is allowed to commit to the low-dimensional target only partially: the overall estimate is the linear mixture of the sample covariance and the target
\begin{equation}
\hat\Sigma_{\hat d,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat  T_{\hat d}
\end{equation}

>>>>

The hyperparameters $\hat\eta$ and $\hat\lambda$ must be estimated from the data.  Under the MSE loss function $\mathcal L_e$ (\ref{eq:MSE}), the optimal values of $\hat\eta$ and $\hat\lambda$ can be estimated analytically \cite{Ledoit:2004,Schafer:2005,Schaefer:2010}. However, these estimates are no longer optimal under the Gaussian loss $\mathcal L_g$. \TODO{I have tested this and there was a substantial difference in both synthetic and empirical data}  When an analytical solution for optimal hyperparameters is not available, the optimal values can be found by cross validation, entirely within the the training sample. \TODO{explain nested cross-validation in more detail?} \TODO{KJ: Yes - see above.}

\TODO{KJ: I am confused.  This suggests that you are getting different results using the two different loss functions.
However, there was no argument yet that either gives us more relevant information about the underlying distribution
if we do not assume Gaussianity, for instance.  I would argue that the Gaussian loss is better, but this is simply 
because the Euclidean metric is not appropriate on the cone of positive semidefinite variances.  As the 
discussion we have been having with Alex demonstrates, this is a point that needs to be explained clearly.}

@@

\subsection*{Estimator B: Shrinkage toward a factor model}
For estimator B, the target is a factor model, composed as the sum of a low-rank component $\hat L_{\hat d}\hat L_{\hat d}^\T$ and a diagonal matrix $\hat \Psi$:
\begin{equation}
\hat T_{\hat d} = \hat L_{\hat d} \hat L_{\hat d}^\T + \hat\Psi
\end{equation}
Here $\hat L_{\hat d}$ is an $p\times\hat d$ matrix and $\hat\Psi$ is diagonal.

@@

Just as with Estimator A, Estimator B is allowed to commit to the low-dimensional target only partially: the overall estimate is the linear mixture of the sample covariance and the target
\begin{equation}
\hat\Sigma_{\hat d,\hat\lambda} = (1-\hat\lambda)\hat\Sigma_0 + \hat  T_{\hat d}
\end{equation}

### 2013-09-23 sections/Theory.tex
minor corrections
Kreso's comments

\subsection*{Basic concepts}
Let $x(t) \in \mathcal X, t=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $t$.  
The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
For example, in our applicaiton, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x(t)$ to take on negative and non-integer values.

We do not assume that the observations $x(t)$ are independently distributed. Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distirbution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.

@@

And, for an ergodic process, $\max\limits_{x\in\mathcal X} \left|\hat F_n(x) - F(x)\right| \to 0$ as $n$ increases.

@@

But the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
\begin{equation}
\hat\Sigma_0 =  \frac 1 {n-c} \sum\limits_{t=1}^n (x(t) -\hat\mu) (x(t) - \hat\mu)^\T
\quad\mbox{where}\;
\hat\mu = \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $c$ is the sum of temporal correlations \TODO{there must be a better description of this}. For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
\TODO{Under the Gaussian loss, $\mathcal L_g$ (\ref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.  But it would be important if we used $\mathcal L_e$ (\ref{eq:MSE}).}

@@

The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions. When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{t=1}^n x(t) x(t)^\T$ is  
\begin{equation}
L\left(\Sigma \mid \hat\Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
\end{equation}
Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
\begin{equation}\label{eq:GaussLoss}
\mathcal L_g(\hat\Sigma,\Sigma) 
=  -\frac 2 {pn} L\left(\Sigma \mid \hat\Sigma \right) - \frac 1 p \ln(2\pi) 
\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}) \right) 
\end{equation}
The corresponding excess loss 
\begin{equation}
\ell_g(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) - \mathcal L_g(\Sigma,\Sigma)  
= \frac 1 p \left(-\ln \det (\hat \Sigma^{-1} \Sigma) + \Tr(\hat \Sigma^{-1}\Sigma)\right) - 1
\end{equation}
is known as \emph{entropy loss} \cite{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}

>>>>

\subsection*{Basic concepts}
Let $x(t) \in \mathcal X, t=1\ldots,n$ denote a sample of consecutive observations of population activity of $p$ neurons in time bins $t$.  
The activity of each neuron is denoted by the real-valued firing rate, thus  $\mathcal X = \mathbb R^{p\times 1}$.  
For example, in our application, the firing rate is estimated from somatic calcium fluorescence signals with the mean stimulus response subtracted, making it possible for elements of $x(t)$ to take on negative and non-integer values.

We do not assume that the observations $x(t)$ are independently distributed. Rather, the data generating process is assumed to be \emph{ergodic}, \emph{i.e.}\;described by a \emph{true distribution} $F: \mathcal X \mapsto  \mathbb [0, 1]$ (cumulative) over long periods of time such that the \emph{empirical distribution} $\hat F_n$ from a given sample will convege to $F$ with increasing sample size $n$.

@@

For an ergodic process, $\max\limits_{x\in\mathcal X} \left|\hat F_n(x) - F(x)\right| \to 0$ as $n$ increases.  \TODO{If you decide to leave this in, provide a reference to a standard book in the field.}

@@

But the usual estimator of the true covariance matrix $\Sigma$ is  the sample covariance matrix:
\begin{equation}
\hat\Sigma_0 =  \frac 1 {n-c} \sum\limits_{t=1}^n (x(t) -\hat\mu) (x(t) - \hat\mu)^\T
\quad\mbox{where}\;
\hat\mu = \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $c$ is the sum of temporal correlations \TODO{there must be a better description of this} \TODO{Which?  I think the equation is OK.  Do you mean $c$?  This is standard, and again, I suggest just providing a reference.}. For independent observations, $c=1$;  $c>1$ when nearby samples are correlated; the value of $c$ can be estimated from the data.
\TODO{Under the Gaussian loss, $\mathcal L_g$ (\ref{eq:GaussLoss}) the scaling is irrelevant, so we will not spend any time talking about the scaling.  But it would be important if we used $\mathcal L_e$ (\ref{eq:MSE}).}

@@

The Gaussian loss function $\mathcal L_g$  arises from the theory of multivariate normal distributions. When observations are identically and independently distributed according to a multivariate normal distribution with zero means, the log likelihood of the covariance matrix $\Sigma$ with $\hat\Sigma = \frac 1 n \sum\limits_{t=1}^n x(t) x(t)^\T$ is  
\begin{equation}
L\left(\Sigma \mid \Sigma\right) = -\frac n 2 \ln(2\pi) - \frac n 2 \ln \det \Sigma - \frac n 2 \Tr(\Sigma^{-1} \hat \Sigma)
\end{equation}
Then $\mathcal L_g$ is constructed by rescaling $L\left(\Sigma \mid \hat\Sigma\right)$ and dropping the constant term:
\begin{equation}\label{eq:GaussLoss}
\mathcal L_g(\hat\Sigma,\Sigma) 
=  -\frac 2 {pn} L\left(\hat\Sigma \mid \Sigma \right) - \frac 1 p \ln(2\pi) 
\equiv  \frac 1 p\left(\ln \det \hat \Sigma + \Tr(\hat \Sigma^{-1}\Sigma) \right) 
\end{equation}
\TODO{Note that the switch from $L\left(\Sigma \mid \hat\Sigma \right)$ to $L\left(\hat\Sigma \mid \Sigma \right)$ is on purpose, which will become clear in the next few steps.}
The corresponding excess loss 
\begin{equation}
\ell_g(\hat\Sigma,\Sigma) = \mathcal L_g(\hat\Sigma,\Sigma) - \mathcal L_g(\Sigma,\Sigma)  
= \frac 1 p \left(-\ln \det (\hat \Sigma^{-1} \Sigma) + \Tr(\hat \Sigma^{-1}\Sigma)\right) - 1
\end{equation}
is known as \emph{entropy loss} \cite{James:1961}. \TODO{The order of the operands is reversed. Check that it's okay.}
